---
title: "The Properties of the Exponential Distribution"
author: "By: Mike Silva"
date: "September 2014"
output:
  pdf_document: default
  html_document:
    theme: cerulean
---

To explore the properties of exponential distributions we will conduct a simulation.  Each simulation will randomly select 40 exponential distributions and average them.  The simulation will run 1,000 times.  Once simulated we will explore the properties based on the sample and make inferences.  

## Simulation of Exponential Distribution  
```{r, echo=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
```

```{r}
## Simulation Parameters 
lambda <- 0.2
simulations <- 1000
number.of.exponentials.in.a.mean <- 40

## Objects to hold the results
sim.results <- mean.results <- data.frame(value=numeric())

## Simulation loop
for(i in 1:simulations) {
  ## Run the Simulation
  simulation <- rexp(number.of.exponentials.in.a.mean, lambda)
  ## Load simulation results into sim.results data frame
  sim.results <- rbind_list(sim.results, data.frame(value = simulation))
  ## Compute the mean of the simulations and load it into the mean.results data frame
  mean.results <- rbind_list(mean.results, data.frame(value = mean(simulation)))
}
```

## Results  
Now that we have completed the simulation let's visualize the results.  First let's explore the distribution of the randomly selected observations:  

```{r, echo=FALSE}
ggplot(sim.results, aes(x = value)) + geom_histogram(binwidth=1, aes(y = ..density..), fill='#FF9999', color='black') + ggtitle(paste('Figure 1: Distribution of the Result of the', prettyNum(simulations, big.mark=','),'Simulations')) + theme(axis.title.x = element_blank()) + ylab('Density')
```

We observe they are indeed an exponential distribution.  Now let's look at the value of the mean of `r number.of.exponentials.in.a.mean` exponentials:  

```{r, echo=FALSE}
ggplot(mean.results, aes(x = value)) + geom_histogram(binwidth=0.2, aes(y = ..density..), fill='#FF9999', color='black') + ggtitle(paste('Figure 2: Distribution of the Means of',number.of.exponentials.in.a.mean,'Simulations')) + geom_vline(xintercept = mean(mean.results$value), color='red', linetype = "longdash") + theme(axis.title.x = element_blank()) + ylab('Density') + stat_function(fun=dnorm,colour = "#009E73",arg = list(mean = 1/lambda,sd = 1/lambda*(1/sqrt(number.of.exponentials.in.a.mean))))
```

The following observations can be made about exponential distributions:  

*  The average, shown as the red line in figure 2 is `r mean(sim.results$value)` which is a close aproximation of the theoretical mean of `r 1/lambda`.  
*  The variance of the simulation is `r sd(sim.results$value)^2` which is again a close appoximation of the theoretical variance of `r (1/lambda)^2`.  The standard deviation of the simulation is `r sd(sim.results$value)` which is close to `r 1/lambda`.  
*  We observe that the distribution of the means fit in the green normal curve.  
*  We observe that % of the observations fall 
