---
title: "Tooth Growth"
author: "By: Mike Silva"
date: "September 2014"
output:
  pdf_document: default
  html_document:
    theme: cerulean
---
```{r, echo=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(pander)
```
## Introduction

The response is the length of teeth in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).

## Exploratory Data Analyses  
We begin by looking at the relationship between tooth length, dose, and delivery method.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=8}
## Visualize the data
ggplot(data = ToothGrowth, aes(x=dose, y=len, color=supp))+geom_point()+xlab('Dose in Milligrams')+ylab('Tooth Length')+facet_grid(. ~ supp)+theme(legend.position='none')+ggtitle('Figure 1: Tooth Length by Dose and Delivery Method (O.J. or Vitamin C)')+stat_smooth()
```

**Table 1: Tooth Length by Dose and Delivery Method**  
```{r, results='asis', echo=FALSE}
data <- ToothGrowth %.%
  group_by(supp, dose) %.%
  summarize(m.length = mean(len),
            sd.length = sd(len))
names(data)<-c('Method','Dose (mg)','Mean','Standard Deviation')

pandoc.table(data, big.mark = ',')
```
We observe the tooth length the relationship between the dose and tooth length is positive

## Comparison of Tooth Growth
Given that all the guinea pigs were drawn from the same population and the population variance is the same, I assumed the sample variance is constant.  

### By Suppliment
```{r}
oj <- filter(ToothGrowth, supp == 'OJ') %.% select(len)
vc <- filter(ToothGrowth, supp == 'VC') %.% select(len)
t.test(oj,vc)
```

### By Dose
Text
