# The Properties of the Exponential Distribution 
**Author: Mike Silva <<mike.a.silva@gmail.com>>**  
**Last Ran: `r Sys.time() `**  

## Simulation Parameters  

The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also also 1/lambda. Set lambda = 0.2 for all of the simulations. In this simulation, you will investigate the distribution of averages of 40 exponential(0.2)s. Note that you will need to do a thousand or so simulated averages of 40 exponentials.

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponential(0.2)s.  You should
1.Show where the distribution is centered at and compare it to the theoretical center of the distribution.
2.Show how variable it is and compare it to the theoretical variance of the distribution.
3.Show that the distribution is approximately normal.
4.Evaluate the coverage of the confidence interval for 1/lambda: X¯±1.96Sn???.

```{r, echo=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
## Simulation Parameters
lambda <- 0.2
simulations.ran <- 3000
number.of.exponentials <- 40
## Objects to hold the results
sim.results <- mean.results <- data.frame(simulation=numeric(), value=numeric(), measure=factor())

## Simulation loop
for(i in 1:simulations.ran) {
  ## Run the Simulation
  simulation <- rexp(number.of.exponentials, lambda)
  ## Load simulation results into sim.results data frame
  results <- data.frame(simulation = rep(i, number.of.exponentials), value = simulation, measure=factor('Simulation'))
  sim.results <- rbind_list(sim.results, results)
  ## Compute the mean of the simulation
  results <- data.frame(simulation = i, value = mean(simulation), measure=factor('Mean'))
  mean.results <- rbind_list(mean.results, results)
  
}

ggplot(sim.results, aes(x = value)) + geom_histogram(binwidth=1, aes(y = ..density..), fill='gray', color='black') + ggtitle(paste('Figure 1: Result of the', prettyNum(simulations.ran, big.mark=','),'Simulations')) + theme(axis.title.x = element_blank()) + ylab('Density')

ggplot(mean.results, aes(x = value)) + geom_histogram(binwidth=0.2, aes(y = ..density..), fill='gray', color='black') + ggtitle('Figure 2: Means of the Simulations') + geom_vline(xintercept = mean(mean.results$value), color='red', linetype = "longdash") + theme(axis.title.x = element_blank()) + ylab('Density')

```

## Properties of the Distribution  

### Center of the Distribution

The average, shown as the red line in the figure above is `r mean(results)` which is a close aproximation of the population `r 1/lambda`. 

### Variance of the Distribution  

The variance of the simulation is `r sd(results)^2` which is again a close appoximation of the population variance of `r 1/lambda`.  The standard deviation of the simulation is `r sd(results)` which is close to `r 1/lambda`.

## Appendix  

### Code Sample  

```{r, eval=FALSE}
lambda <- 0.2
simulations.ran <- 1000
number.of.exponentials <- 40
## Run the Simulation
results <- vector()
for(i in 1:simulations.ran) {
  results[i] <- mean(rexp(number.of.exponentials, lambda))
}
```
```{r, echo=FALSE}
#results2 <- apply(matrix(rexp(number.of.exponentials, lambda), simulations.ran), 1, mean)
```
