---
title: "The Properties of the Exponential Distribution"
author: "Mike Silva"
date: "September 2014"
output:
  html_document: default
  pdf_document: default
---

## Synopsis  
To explore the properties of exponential distributions we conducted a simulation.  In each simulation we randomly selected 40 exponential distributions and average them.  The simulation was run 1,000 times.  Once simulated we explored the properties of the sample.  

## Simulation of Exponential Distribution  
```{r, message=FALSE}
library(dplyr)
## Simulation Parameters 
lambda <- 0.2
simulations <- 1000
number.of.exponentials.in.a.mean <- 40

## Objects to hold the results
sim.results <- mean.results <- data.frame(value=numeric())

## Simulation loop
for(i in 1:simulations) {
  ## Run the Simulation
  simulation <- rexp(number.of.exponentials.in.a.mean, lambda)
  ## Load simulation results into sim.results data frame
  sim.results <- rbind_list(sim.results, data.frame(value = simulation))
  ## Compute the mean of the simulations and load it into the mean.results data frame
  mean.results <- rbind_list(mean.results, data.frame(value = mean(simulation)))
}
```

## Results  
Now that we have completed the simulation let's visualize the results.  First let's explore the distribution of the randomly selected observations:  
```{r, fig.width=7, fig.height=4, echo=FALSE}
library(ggplot2)
ggplot(sim.results, aes(x = value)) + geom_histogram(binwidth=1, aes(y = ..density..), fill='#FF9999', color='black') + ggtitle(paste('Figure 1: Distribution of the Result of the', prettyNum(simulations, big.mark=','),'Simulations')) + theme(axis.title.x = element_blank()) + ylab('Density')
```

We observe the simulation is indeed an exponential distribution.  Now let's look at the distribution of the mean of `r number.of.exponentials.in.a.mean` exponentials:  

```{r, echo=FALSE}
mu <- 1/lambda
ci.lower.bound <-mu - ((1.96*(1/lambda))/sqrt(number.of.exponentials.in.a.mean))
ci.upper.bound <-mu + ((1.96*(1/lambda))/sqrt(number.of.exponentials.in.a.mean))
ggplot(mean.results, aes(x = value)) + geom_histogram(binwidth=0.2, aes(y = ..density..), fill='#FF9999', color='black') + ggtitle(paste('Figure 2: Distribution of the Means of',number.of.exponentials.in.a.mean,'Simulations')) + geom_vline(xintercept = mean(mean.results$value), color='red', linetype = "longdash") + theme(axis.title.x = element_blank()) + ylab('Density') + stat_function(fun=dnorm,colour = "#009E73",arg = list(mean = 1/lambda,sd = 1/lambda*(1/sqrt(number.of.exponentials.in.a.mean)))) + geom_vline(xintercept = ci.lower.bound, color='blue', linetype = "longdash") + geom_vline(xintercept = ci.upper.bound, color='blue', linetype = "longdash")
```

The following observations can be made about exponential distributions:  

*  The average (shown as the red line in figure 2) is `r mean(sim.results$value)` which is a close aproximation of the theoretical mean of `r 1/lambda`.  
*  The variance of the simulation is `r sd(sim.results$value)^2` which is again a close appoximation of the theoretical variance of `r (1/lambda)^2`.
*  The standard deviation of the simulation is `r sd(sim.results$value)` which is close to `r 1/lambda`.  
*  We observe that the distribution of the means generally fits in the green normal curve.  
*  We also observe that `r (length(mean.results[mean.results$value > ci.lower.bound & mean.results$value < ci.upper.bound,]) / nrow(mean.results)) * 100`% of the means fall within the 95% confidence interval (the blue dashed lines in figure 2).
